Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-07-19 article_springer/article_springer.Rmd
Address Referee 1's comments

Often a magnitude, or mark can be assigned to each event in the renewal process, 
such as for earthquakes, solar flares or neuron voltages. 
The Peaks-Over-Threshold model [POT, see e.g. @ColesBook] applies a
threshold to the magnitudes, and fits a Generalized Pareto distribution 
to the threshold exceedances. 
A commonly made assumption in POT models is that times between 
events are either fixed 
or light-tailed, and this entails that the 
threshold crossing times form a Poisson process [@Hsing88]. 
Then as one increases the threshold $\ell$ and thus decreases the threshold 
crossing probability $p_{\ell}$, the Poisson process is rarefied, i.e. its 
intensity decreases _linearly_ with $p_{\ell}$ [see e.g. @beirlantBook].

As will be shown below, in the heavy-tailed waiting time scenario threshold 
crossing times form a _fractional Poisson process_ 
[@Laskin2003; @Meerschaert2010b], which is a 
renewal process with Mittag-Leffler distributed waiting times. 
The family of Mittag-Leffler distributions nests the exponential 
distribution [@Haubold11], and hence the fractional Poisson process
generalizes the standard Poisson process. 
Again as the threshold size $\ell$ increases and the threshold crossing
probability $p_{\ell}$ decreases, the fractional Poisson process is rarefied: 
The scale parameter of the Mittag-Leffler inter-arrival times of 
threshold crossing times increases, but _superlinearly_; 
see the Theorem below.

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
We assume for the rest of the article, that the magnitudes $(J_i)_{i \in \mathbb{N}}$ belong to the the max-domain of attraction of some non-degenerate distribution. This means there exist $a_n>0$ and $d_n \in \mathbb{R}$ such that 
\begin{align} \label{assumptionJs}
a_n^{-1}(J_1 \vee \ldots \vee J_n -d_n) \Rightarrow A \text{ as } n \rightarrow \infty.
\end{align}
Hence, the distribution of $A$ is a generalized extreme value distribution (GEV) which distribution function is given by 
$$ F(x;\xi) = \begin{cases}\exp(-(1+\xi x)^{-1/\xi}) & \xi\neq0 \\ \exp(-\exp(-x)) & \xi = 0\end{cases}.$$ 
The GEV is subdivided into the Gumbel ($\xi=0$), the Weibull ($\xi<0$) and the Fréchet ($\xi>0$) family of distributions.

@@

In this section we state and prove the key theorem. For an accessible introduction to regular variation and stable limit theorems, we recommend the book by @MeerschaertSikorskii.

@@

: For the magnitudes $J_k$ assumption \eqref{assumptionJs} holds. Furthermore, let the waiting times $W_k$ be in the domain of attraction of a positively skewed sum-stable law with stability parameter $0 < \beta < 1$; more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. Write $p_{\ell} := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p_{\ell})} \to Z_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $Z_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sZ_\beta)] = \frac{1}{1+s^\beta}.
$$

*Proof of Theorem:* 	Due to assumption \eqref{assumptionJs}, it follows
	\begin{align*}
	P\left(\frac{M(c)-d(c)}{a(c)} \leq x\right) \rightarrow F(x) \text{ as } c \rightarrow \infty,
	\end{align*}
	where $x$ is any value from the support of $F$. Then it follows,
	\begin{align*}
	  F(xa(c)+d(c))^{\lfloor c \rfloor} &\rightarrow F(x) \text{ as } c \rightarrow \infty,\\
      \Rightarrow  \lfloor c \rfloor \log F(\ell_c)  &\rightarrow \log F(x)\text{ as } c \rightarrow \infty,
	\end{align*}
	with $\ell_c:=xa(c)+d(c)$. Furthermore, since $\log(1-x) \sim -x$ for small $x$ it follows
    \begin{align}
      c \cdot p(\ell_c) \rightarrow -\log F_A(x)  \text{ as } c \rightarrow \infty,\label{eq:lplconv}
    \end{align}
	 with $p(\ell_c):= 1-F(\ell_c)$. Since $\tau(\ell_c) \sim Geo(p(\ell_c))$ and \eqref{eq:lplconv}
	it follows that $\tau(\ell_c)/c$ converges to an Exponential distributed random variable,
	\begin{align*}
	\frac{\tau(\ell_c)}{c} \Rightarrow E  \text{ as } c \rightarrow \infty
	\end{align*}
	 with inverse mean $\lambda:=-\log F_A(x)$. Due to
	\begin{align*}
	S(c):= \sum_{i=1}^{\lfloor c \rfloor} \frac{W_i}{b(c)} \Rightarrow D \ \text{ as } c \rightarrow \infty,
	\end{align*}
    it follows with Gnedenkos transfer theorem (see @gnedenko1983limit), that
	\begin{align}
	\sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(\ell_c)}  \Rightarrow Z\ \text{ as }  c \rightarrow \infty, \label{TheoGutArg}
	\end{align}
	where the distribution of $Z$ has the characteristic function
	\begin{align}
    \Psi(s) &= \int_0^{\infty} (\Phi_D(s))^{y} F_E(dy) \nonumber\\
    &= \int_0^{\infty} \frac{1}{1-\log(\Phi_D(s))/\lambda} \nonumber\\
    &= \frac{1}{1+s^{\beta}\lambda^{-1}}
	\end{align}
    where $\Phi_D(s) = \exp(-s^\beta)$ is the Laplace transform of $D$ 
    and  $F_E(dy)$ is the distribution function of $E$. Hence, $Z$ is Mittag-Leffler distributed with parameter $\lambda^{-1/\beta}$. Rewriting
	\begin{align*}
    \sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(1/p(\ell_c))} = \left(\sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(c) }\right) \frac{b(c)}{b(1/p(\ell_c))}  
	\end{align*}
	and
	\begin{align*}
	\frac{b(c)}{b(1/p(\ell_c))} = \frac{b(c)}{b(c/(p(\ell_c)c))} \sim (p(\ell_c) \cdot c)^{1/\beta} \rightarrow \lambda^{1/\beta} \text{ as } c \rightarrow \infty,
	\end{align*}
	the second factor converges to $\lambda^{1/\beta}$, and it follows that
	 $Z \sim ML(\beta,1)$. Since $c \rightarrow \infty$ is equivalent to $\ell \uparrow x_R$, the assertion follows with $\ell:=\ell_c$ and $p_\ell:=p(\ell_c)$.
\qed

@@

Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the R Package `MittagLeffleR` [@MittagLeffleR], Maximum Likelihood estimation is implemented via numerical optimization. The MLE slightly outperforms the log-Moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computationally intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes. For smaller tails both estimators show an increasing rmse for the scale estimation due to an increasing variance. This results from the fact that in case of very small tails, single very large values can occur.

```{r MSE, echo=FALSE,warning=FALSE, fig.height=7,fig.width=8,fig.align="center", fig.cap="\\label{fig:MSE} Rmse for the estimation of tail (left column) and scale (right column) parameters via log-Moment estimator and MLE of a Mittag-Leffler sample with varying sample size n=25, 100, 200, varying tails on the x-axis and fixed scale equal to one, based on 5000 simulation runs.", out.width = '90%'}
load(here::here("article_springer", "article_springer_files", "ML_estim_simu","tbl_ML_estim.RData"))
  n.labs <-c("n=25", "n=100", "n=200")
  names( n.labs ) <-  c("25", "100", "200")
tbl_ML_estim %>% filter(method=="logMom" | method=="ML")%>%
  filter(n == "25" || n == "100" || n=="200") %>%
  filter(true.scale == "1") %>% 
  ggplot(aes(true.tail, rmse, colour = method, group=method)) + 
  geom_line() + geom_point() + 
  facet_wrap(c("n","param"), scales = "free", ncol = 2, labeller = labeller(n = n.labs)) + 
  scale_color_manual(values=c("red", "blue"),name = "Method", labels = c("log-Moment", "MLE")) + ggtitle("Rmse for log-Moment and MLE") +
   theme(plot.title = element_text(hjust = 0.5))

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. The Hill estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. It heavily depends on the choice of $r$ in \eqref{eq:Hill}, and hence on the number of order statistics entering \eqref{eq:Hill}.  Figure \ref{fig:Hillplots} shows 100 Hill plots (grey thin lines) for simulated Mittag-Leffler data with tail parameter $\beta=0.8$ and scale $\sigma=1$, for a sample size of $n=200$ (left) and $n=1000$ (right). The dark lines are the means and the red dotted lines show the true tail
parameter. To deduce the correct tail parameter estimates from one of those Hillplots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a small tuning parameter $r$. In case that only events above a high threshold are recorded, $n=10000$ events are extremely unrealistic. There is a vast literature on bias-reduced versions of the Hill estimator and on the optimal choice of $r$, mainly based on imposed second-order regular variation conditions [see e.g. @beirlantBook, Ch.4.4 and 4.7 and references in it]. Also, there are hundreds other tail-index estimators not considered in this article, since this would go beyond the scope of our article. For example the moment estimator [@dekkers1989moment], the Pickands estimator [@pickands1975statistical], the QQ-estimator [@Kratz96], to name a few. A nice overview can be found in [@fedotenkov2018review]. However, we know from the Theorem in Section \ref{sec:scaling} that the inter-exceedance times are asymptotically Mittag-Leffler distributed, and hence it is unsurprising that general tail (and scale) estimation methods that do not take into account this structure are statistically less efficient [@kozubowski2001]. Finally, the Hill estimator completely fails in the important special case where the inter-arrival times are exponentially distributed and not heavy-tailed. For these reasons, we propose to use the log-Moment estimator or the MLE for Mittag-Leffler distributions.

```{r Hillplots,  fig.height=5, fig.width=4, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for m=100 simulated Mittag-Leffler datasets with true tail 0.8 and sample size 200 (first panel) and 10000 (second panel), with number of upper order statistics r on which the Hill estimator is based on the x-axis. The grey thin lines are the Hill plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. ", out.width = '50%'}

@@

grid.arrange(p1,p2,nrow=2,top="Hillplots") 
```

@@

We have extended the POT (Peaks-over-Threshold) model, a mainstay of extreme value theory, to "bursty" time series, which have been studied 
intensively in statistical physics. Burstiness is characterized by power-law waiting times between events, and we have shown that the Mittag-Leffler distribution arises naturally as a scaling limit for the inter-exceedance times of high thresholds. Moreover, we have derived the following non-linear scaling behaviour: $\sigma \sim p_{\ell}^{-1/\beta}$, where $\sigma$ is the scale parameter of the distribution of threshold 
exceedance times, $p_{\ell}$ is the fraction of magnitudes above the threshold $\ell$, and $\beta$ the exponent of the power law. 
This "anomalous" scaling behaviour in the bursty setting entails two phenomena:

@@

The bursty CTRE approach to model "non-Poissonian" threshold crossing times should be contrasted with the (now standard) approach of clusters of extremes, see e.g. @ferro2003inference. In this approach, i.i.d. event sequences of magnitudes are generalized to stationary sequences of
event magnitudes (subject to a mixing condition). The two approaches are fundamentally different: Clustering assumes that each event belongs to one particular (latent) group of events. For bursts, however, the aim is to identify an underlying scale-free pattern in the event 
dynamics, which is often characteristic of complex systems. It is an interesting open problem to develop criteria, based e.g. on measures of surprise [@Lee15], that guide an applied statistician in the choice between a clustering and a CTRE approach for a particular problem. 
In sum, the CTRE approach should not be seen as a substitute for the usual cluster approach, but rather as a complementary approach.

Moreover, it was observed in @towe2020modelling that clusters of extreme events still exist after declustering. We believe it is possible to marry the clustering approach with our CTRE approach by considering CTREs based on MRPs with a _stationary_, rather than i.i.d., sequence of magnitudes, so that the times between clusters are asymptotically Mittag-Leffler distributed. More broadly speaking, the identification of clusters in the context of heavy-tailed inter-arrival times is an open problem.

Finally, a purely scale-free pattern for event times may be too rigid an 
assumption for some bursty time series, because often the heavy-tailed character of the inter-arrival time
distribution does not hold at all time scales. Rather, it applies at short and intermediate time scales, and is truncated (or tempered, 
reverting to an exponential distribution) at very long time scales [see e.g. @MeerschaertRoyQin; and @Aban06]. In such situations, a "tempered" Mittag-Leffler distribution may provide a more realistic fit, which we aim to introduce in follow-up work.

>>>>

Often a magnitude, or mark can be assigned to each event in the renewal process, 
such as for earthquakes, solar flares or neuron voltages. 
The Peaks-Over-Threshold model [POT, see e.g. @ColesBook] applies a
threshold to the magnitudes, and fits a Generalized Pareto distribution 
to the threshold exceedances. 
A commonly made assumption in POT models is that times between 
events are either fixed 
or light-tailed, and this entails that the 
threshold crossing times form a Poisson process [@Hsing88]. 
Then as one increases the threshold $\ell$ and thus decreases the threshold 
crossing probability $p_{\ell}$, the Poisson process is thinned, i.e. its 
intensity decreases _linearly_ with $p_{\ell}$ [see e.g. @beirlantBook].

As will be shown below, in the heavy-tailed waiting time scenario threshold 
crossing times form a _fractional Poisson process_ 
[@Laskin2003; @Meerschaert2010b], which is a 
renewal process with Mittag-Leffler distributed waiting times. 
The family of Mittag-Leffler distributions nests the exponential 
distribution [@Haubold11], and hence the fractional Poisson process
generalizes the standard Poisson process. 
Again as the threshold size $\ell$ increases and the threshold crossing
probability $p_{\ell}$ decreases, the fractional Poisson process is thinned: 
The scale parameter of the Mittag-Leffler inter-arrival times of 
threshold crossing times increases, but _superlinearly_; 
see the Theorem below.

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
We assume for the rest of the article, that the magnitudes $(J_i)_{i \in \mathbb{N}}$ belong to the the max-domain of attraction of some non-degenerate distribution. This means there exist $a_n>0$ and $d_n \in \mathbb{R}$ such that 
\begin{align} \label{assumptionJs}
a_n^{-1}(J_1 \vee \ldots \vee J_n -d_n) \Rightarrow A \text{ as } n \rightarrow \infty.
\end{align}
Hence, the distribution of $A$ is a generalized extreme value distribution (GEV) whose distribution function is given by 
$$ F(x;\xi) = \begin{cases}\exp(-(1+\xi x)_+^{-1/\xi}) & \xi\neq0 \\ \exp(-\exp(-x)) & \xi = 0\end{cases}$$ 
where $(.)_+:= \max\{0,.\}$.

The GEV is subdivided into the Gumbel ($\xi=0$), the Weibull ($\xi<0$) and the Fréchet ($\xi>0$) family of distributions.

@@

In this section we state and prove the key theorem, which is founded on the concept of regular variation. A Borel-measurable function $f:\mathbb (0,\infty) \to (0,\infty)$ is said be "regularly varying at $\infty$ with parameter (or "index") $\rho$ if

$$
\frac{f(\lambda x)}{f(x)} = \lambda^\rho \quad \text{ for all } \lambda > 0.
$$

For more details regular variation and stable limit theorems, we recommend the book by @MeerschaertSikorskii.

@@

: For the magnitudes $J_k$ assumption \eqref{assumptionJs} holds. Furthermore, let the waiting times $W_k$ be in the domain of attraction of a positively skewed sum-stable law with stability parameter $0 < \beta < 1$; more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$, $s > 0$. Write $p_{\ell} := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p_{\ell})} \to Z_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $Z_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sZ_\beta)] = \frac{1}{1+s^\beta}.
$$

*Proof of Theorem:* 	Due to assumption \eqref{assumptionJs}, it follows that
	\begin{align*}
	P\left(\frac{M(c)-d(c)}{a(c)} \leq x\right) \rightarrow F(x) := F(x,\xi) \text{ as } c \rightarrow \infty,
	\end{align*}
	where $x$ is any value from the support of $F$. Then
	\begin{align*}
	  F(xa(c)+d(c))^{\lfloor c \rfloor} &\rightarrow F(x) \text{ as } c \rightarrow \infty,\\
      \Rightarrow  \lfloor c \rfloor \log F(\ell_c)  &\rightarrow \log F(x)\text{ as } c \rightarrow \infty,
	\end{align*}
	with $\ell_c:=xa(c)+d(c)$. Furthermore, since $\log(1-x) \sim -x$ for small $x$ it follows
    \begin{align}
      c \cdot p(\ell_c) \rightarrow -\log F_A(x)  \text{ as } c \rightarrow \infty,\label{eq:lplconv}
    \end{align}
	 with $p(\ell_c):= 1-F(\ell_c)$. Since $\tau(\ell_c) \sim Geo(p(\ell_c))$ and \eqref{eq:lplconv}
	it follows that $\tau(\ell_c)/c$ converges to an Exponential distributed random variable,
	\begin{align*}
	\frac{\tau(\ell_c)}{c} \Rightarrow E  \text{ as } c \rightarrow \infty
	\end{align*}
	 with inverse mean $\lambda:=-\log F_A(x)$. Due to
	\begin{align*}
	S(c):= \sum_{i=1}^{\lfloor c \rfloor} \frac{W_i}{b(c)} \Rightarrow D \ \text{ as } c \rightarrow \infty,
	\end{align*}
    it follows with Gnedenko's transfer theorem (see @gnedenko1983limit), that
	\begin{align}
	\sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(\ell_c)}  \Rightarrow Z\ \text{ as }  c \rightarrow \infty, \label{TheoGutArg}
	\end{align}
	where the distribution of $Z$ has the characteristic function
	\begin{align}
    \Psi(s) &= \int_0^{\infty} (\Phi_D(s))^{y} F_E(dy) \nonumber\\
    &= \int_0^{\infty} \frac{1}{1-\log(\Phi_D(s))/\lambda} \nonumber\\
    &= \frac{1}{1+s^{\beta}\lambda^{-1}}
	\end{align}
    where $\Phi_D(s) = \exp(-s^\beta)$ is the Laplace transform of $D$ 
    and  $F_E(dy)$ is the distribution function of $E$. Hence, $Z$ is Mittag-Leffler distributed with parameter $\lambda^{-1/\beta}$. Rewriting
	\begin{align*}
    \sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(1/p(\ell_c))} = \left(\sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(c) }\right) \frac{b(c)}{b(1/p(\ell_c))}  
	\end{align*}
	and
	\begin{align*}
	\frac{b(c)}{b(1/p(\ell_c))} = \frac{b(c)}{b(c/(p(\ell_c)c))} \sim (p(\ell_c) \cdot c)^{1/\beta} \rightarrow \lambda^{1/\beta} \text{ as } c \rightarrow \infty,
	\end{align*}
	the second factor converges to $\lambda^{1/\beta}$, and it follows that
	 $Z \sim ML(\beta,1)$. Since $c \rightarrow \infty$ is equivalent to $\ell \uparrow x_R$, the assertion follows with $\ell:=\ell_c$ and $p_\ell:=p(\ell_c)$.
\qed

@@

Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the R Package `MittagLeffleR` [@MittagLeffleR], Maximum Likelihood estimation is implemented via numerical optimization. The MLE slightly outperforms the log-Moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computationally intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes. For smaller tails both estimators show an increasing RMSE for the scale estimation due to an increasing variance. This results from the fact that in case of very small tails, single very large values can occur.

```{r MSE, echo=FALSE,warning=FALSE, fig.height=7,fig.width=8,fig.align="center", fig.cap="\\label{fig:MSE} RMSE for the estimation of tail (left column) and scale (right column) parameters via log-Moment estimator and MLE of a Mittag-Leffler sample with varying sample size n=25, 100, 200, varying tails on the x-axis and fixed scale equal to one, based on 5000 simulation runs.", out.width = '90%'}
load(here::here("article_springer", "article_springer_files", "ML_estim_simu","tbl_ML_estim.RData"))
  n.labs <-c("n=25", "n=100", "n=200")
  names( n.labs ) <-  c("25", "100", "200")
tbl_ML_estim %>% filter(method=="logMom" | method=="ML")%>%
  filter(n == "25" || n == "100" || n=="200") %>%
  filter(true.scale == "1") %>% 
  ggplot(aes(true.tail, rmse, colour = method, group=method)) + 
  geom_line() + geom_point() + 
  facet_wrap(c("n","param"), scales = "free", ncol = 2, labeller = labeller(n = n.labs)) + 
  scale_color_manual(values=c("red", "blue"),name = "Method", labels = c("log-Moment", "MLE")) + ggtitle("RMSE for log-Moment and MLE") +
   theme(plot.title = element_text(hjust = 0.5))

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. The Hill estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. It heavily depends on the choice of $r$ in \eqref{eq:Hill}, and hence on the number of order statistics entering \eqref{eq:Hill}.  Figure \ref{fig:Hillplots} shows 100 Hill plots (grey thin lines) for simulated Mittag-Leffler data with tail parameter $\beta=0.8$ and scale $\sigma=1$, for a sample size of $n=200$ (left) and $n=1000$ (right). The dark lines are the means and the red dotted lines show the true tail
parameter. To deduce the correct tail parameter estimates from one of those Hill plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a small tuning parameter $r$. In case that only events above a high threshold are recorded, $n=10000$ events are extremely unrealistic. There is a vast literature on bias-reduced versions of the Hill estimator and on the optimal choice of $r$, mainly based on imposed second-order regular variation conditions [see e.g. @beirlantBook, Ch.4.4 and 4.7 and references in it]. Also, there are hundreds other tail-index estimators not considered in this article, since this would go beyond the scope of our article. For example the moment estimator [@dekkers1989moment], the Pickands estimator [@pickands1975statistical], the QQ-estimator [@Kratz96], to name a few. A nice overview can be found in [@fedotenkov2018review]. However, we know from the Theorem in Section \ref{sec:scaling} that the inter-exceedance times are asymptotically Mittag-Leffler distributed, and hence it is unsurprising that general tail (and scale) estimation methods that do not take into account this structure are statistically less efficient [@kozubowski2001]. Finally, the Hill estimator completely fails in the important special case where the inter-arrival times are exponentially distributed and not heavy-tailed. For these reasons, we propose to use the log-Moment estimator or the MLE for Mittag-Leffler distributions.

```{r Hillplots,  fig.height=5, fig.width=4, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hill plots for m=100 simulated Mittag-Leffler datasets with true tail 0.8 and sample size 200 (first panel) and 10000 (second panel), with number of upper order statistics r on which the Hill estimator is based on the x-axis. The grey thin lines are the Hill plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. ", out.width = '50%'}

@@

grid.arrange(p1,p2,nrow=2,top="Hill plots") 
```

@@

We have applied the POT (Peaks-over-Threshold) model, a mainstay of extreme value theory, to "bursty" time series, which have been studied 
intensively in statistical physics. Burstiness is characterized by power-law waiting times between events, and we have shown that the Mittag-Leffler distribution arises naturally as a scaling limit for the inter-exceedance times of high thresholds. Moreover, we have derived the following non-linear scaling behaviour: $\sigma \sim p_{\ell}^{-1/\beta}$, where $\sigma$ is the scale parameter of the distribution of threshold 
exceedance times, $p_{\ell}$ is the fraction of magnitudes above the threshold $\ell$, and $\beta$ the exponent of the power law. 
This "anomalous" scaling behaviour in the bursty setting entails two phenomena:

@@

The CTRE approach to model "non-Poissonian" threshold crossing times should be contrasted with the well-documented approach of clusters of extremes, see e.g. @ferro2003inference. When the underlying stochastic process is *stationary*, the exceedances of high thresholds form, asymptotically, a *Cluster Poisson Process*.
In the setting of this article, however, clustering-like dynamics occur due to the *non-stationarity* of the underlying renewal-reward process, which has infinite-mean renewal times. 
Hence CTRE and Cluster Poisson Process shouldn't be viewed as competing methods, as the underlying data generating processes are quite different.
In weighing the evidence for either of the two data generating processes, criteria could be developed, based e.g. on measures of surprise [@Lee15], which may prove to be valuable for future applied statisticians.

Finally, we note that assuming a purely scale-free pattern for event times may be too rigid an 
assumption, which unnecessarily limits the applicability of CTREs. Often, the heavy-tailed character of the inter-arrival time
distribution holds at short to intermediate time scales, and is truncated (or tempered, 
reverting to an exponential distribution) at very long time scales [see e.g. @MeerschaertRoyQin; and @Aban06]. In such situations, a "tempered" Mittag-Leffler distribution may provide a better fit, which we aim to introduce in follow-up work.
